---
title: "Midterm 1 W24"
author: "Amanda Duong"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code must be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Be sure to add your name to the author header above. 

Your code must knit in order to be considered. If you are stuck and cannot answer a question, then comment out your code and knit the document. You may use your notes, labs, and homework to help you complete this exam. Do not use any other resources- including AI assistance.  

Don't forget to answer any questions that are asked in the prompt!  

Be sure to push your completed midterm to your repository. This exam is worth 30 points.  

## Background
In the data folder, you will find data related to a study on wolf mortality collected by the National Park Service. You should start by reading the `README_NPSwolfdata.pdf` file. This will provide an abstract of the study and an explanation of variables.  

The data are from: Cassidy, Kira et al. (2022). Gray wolf packs and human-caused wolf mortality. [Dryad](https://doi.org/10.5061/dryad.mkkwh713f). 

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the wolves data
In these data, the authors used `NULL` to represent missing values. I am correcting this for you below and using `janitor` to clean the column names.
```{r message=FALSE, warning=FALSE}
wolves <- read.csv("data/NPS_wolfmortalitydata.csv", na = c("NULL")) %>% clean_names()
```

## Questions
Problem 1. (1 point) Let's start with some data exploration. What are the variable (column) names?  
```{r}
names(wolves)
```

Problem 2. (1 point) Use the function of your choice to summarize the data and get an idea of its structure.  
```{r}
summary(wolves)
```

Problem 3. (3 points) Which parks/ reserves are represented in the data? Don't just use the abstract, pull this information from the data.  
```{r}
wolves %>%
  count(park) # show which parks are represented
```

Problem 4. (4 points) Which park has the largest number of wolf packs?
```{r}
wolves %>%
  group_by(park) %>%
  summarize(n_distinct_pack=n_distinct(pack)) %>% # only count number of distinct packs (no repeats)
  arrange(desc(n_distinct_pack)) # arrange in desc order to see largest
```
DENA (Denali National Park and Preserve) had the largest number of wolf packs with 69 distinct wolf packs throughout the years.  

Problem 5. (4 points) Which park has the highest total number of human-caused mortalities `mort_all`?
```{r}
wolves %>%
  select(park, mort_all) %>%
  group_by(park) %>%
  summarize(human_mortalities=sum(mort_all, na.rm=T)) %>% # calculate total number of mortalities and remove any NAs
  arrange(desc(human_mortalities))
```
YUCH (Yukon-Charley Rivers National Preserve) had the highest total number of human-caused mortalities.  

The wolves in [Yellowstone National Park](https://www.nps.gov/yell/learn/nature/wolf-restoration.htm) are an incredible conservation success story. Let's focus our attention on this park.  

Problem 6. (2 points) Create a new object "ynp" that only includes the data from Yellowstone National Park.  
```{r}
names(wolves)
```

```{r}
ynp <- wolves %>%
  select(park, biolyr, pack, packcode, packsize_aug, mort_yn, mort_all, mort_lead, mort_nonlead, reprody1, persisty1) %>%
  filter(park=="YNP")
ynp
```

Problem 7. (3 points) Among the Yellowstone wolf packs, the [Druid Peak Pack](https://www.pbs.org/wnet/nature/in-the-valley-of-the-wolves-the-druid-wolf-pack-story/209/) is one of most famous. What was the average pack size of this pack for the years represented in the data?
```{r}
ynp %>%
  filter(pack=="druid") %>% # filter for the druid peak pack
  group_by(biolyr) %>% # group by each year represeted in the data
  summarise(mean_pack_size=mean(packsize_aug, na.rm=T)) # This is the average pack size by year
```
```{r}
ynp %>%
  filter(pack=="druid") %>% # filter for druid peak pack
  summarise(mean_pack_size=mean(packsize_aug, na.rm=T)) # This is the average pack size throughout all of the years
```

Problem 8. (4 points) Pack dynamics can be hard to predict- even for strong packs like the Druid Peak pack. At which year did the Druid Peak pack have the largest pack size? What do you think happened in 2010?
```{r}
ynp %>%
  filter(pack=="druid") %>% # filter for only druid peak pack
  group_by(biolyr) %>% # group by each year represented in the data
  summarise(max_pack_size=max(packsize_aug, na.rm=T)) %>% # find largest druid pack size
  arrange(desc(max_pack_size)) # arrange in desc order to see what year had the largest pack size
```
The Druid Peak pack had the largest pack size in 2001. In 2010, there may not have been any available recorded data since there was 0 recorded for 2010, but in 2008, the pack size was 21 and in 2009, it was 12.  

Problem 9. (5 points) Among the YNP wolf packs, which one has had the highest overall persistence `persisty1` for the years represented in the data? Look this pack up online and tell me what is unique about its behavior- specifically, what prey animals does this pack specialize on?  
```{r}
ynp %>%
  group_by(pack) %>%
  summarise(overall_persistence=sum(persisty1, na.rm=T)) %>%
  arrange(desc(overall_persistence))
```
The pack with the highest overall persistence throughout the years was the mollies, or Mollie's Pack. Their unique behavior includes hunting bison and regular interactions with bears. Bison are the primary prey option for Mollie's pack after elk have migrated away.  

Problem 10. (3 points) Perform one analysis or exploration of your choice on the `wolves` data. Your answer needs to include at least two lines of code and not be a summary function.  

```{r}
wolves %>%
  select(park, biolyr, mort_lead) %>%
  group_by(park) %>%
  filter(between(biolyr, 2005, 2012)) %>%
  summarize(mean_mortlead=mean(mort_lead, na.rm=T)) %>%
  arrange(desc(mean_mortlead))
```

These are the average numbers of human-caused mortalities of a leader wolf for the years 2005-2012. YUCH had the highest average human-caused mortalities of a leader wolf during the selected time period. 