---
title: "BIS15L_AI_hmwk"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
This homework is a bit less traditional this week. There are some errors in the code below. In addition add notes and titles to the code chunks that explains what the code chunck does. Use AI tools, notes from class, or any other resources to help fix the errors. Feel free to edit any plots to your preference. Don't forget to answer the two questions at the bottom. Upon completion, you should not have to '#' out any of the code chunks.

Push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.   

Again make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  


```{r}
library(tidyverse)
library(janitor)
library(dplyr)
```

```{r}
#install.packages("janitor")
```

1.
```{r}
blood_count <- read_csv("data/blood_count_dataset.csv") %>% clean_names() # clean column names
```

2. 
```{r}
glimpse(blood_count) # changed spelling of glimpse
# view summary of data
```

3. 
```{r}
blood_count %>% 
  group_by(gender) %>% # get min, mean, and max of data by gender
  summarize(min=min(red_blood_cells),
            mean=mean(red_blood_cells), # changed spelling of mean
            max=max(red_blood_cells))
```


4. 
```{r}
options(scipen = 999, digits = 2)

blood_count %>% 
  filter(gender == "Male") %>% # capitalized M in male
  select(3:9) %>% # changed brackets
  summarise_all(mean, na.rm = TRUE)
```

5. 
```{r}
options(scipen = 999, digits = 2)

blood_count %>% 
  filter(gender == "Female") %>% 
  select(3:9) %>% # changed brackets
  summarise_all(mean, na.rm = TRUE)
```


6. 
```{r}
blood_count %>%
  pivot_longer(cols = 3:9, names_to = "Test", values_to = "Values") %>% # removed brackets
  ggplot(aes(x = gender, y = Values, fill = gender)) + # lower case g for gender
  geom_boxplot() +
  facet_wrap(~ Test, scales = "free")+ # added "wrap" to facet
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values")
```

7. 
```{r}
grouped_bc <- blood_count %>%
  mutate(Age_Group = case_when( # fixed spelling of age and added mutate function
    age >= 20 & age <= 29 ~ "20s",
    age >= 30 & age <= 39 ~ "30s",
    age >= 40 & age <= 49 ~ "40s",
    age >= 50 & age <= 59 ~ "50s",
    age >= 60 & age <= 70 ~ "60s")) # there were only 4 people in 70s, so I combined it with 60s

head(grouped_bc)

```

8. 
```{r}
grouped_bc %>% # changed blood_count to grouped_bc 
  pivot_longer(cols = 3:9, names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Age_Group, y = Values, fill = Age_Group)) +
  geom_boxplot() +
  facet_wrap(~ Test, scales = "free") +  # facet by the "Test" variable and added "~"
  labs(title = "Boxplot of Blood Count Tests by Age Group",
       x = "Age Group",
       y = "Values")
```
9. What are two drawbacks and two benefits of AI?
2 drawbacks include AI doesn't always know what you're asking it to do so it may take longer to try and write code that is meant for your analysis; another drawback is that AI could interpret your data incorrectly. 2 benefits include if AI knows what you intend to do, it can be much faster to code and AI can easily fix mistakes if your code is more "generic". 
10. Do you think you will use AI for coding in the future? Why or why not?
I think I will use AI to fix errors in my code such as syntax that I could easily miss. 

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 