---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI, or Artificial Intelligence, refers to the development of computer systems that can perform tasks that typically require human intelligence. In other words, it involves creating machines or software that can mimic cognitive functions such as learning, problem-solving, understanding natural language, and recognizing patterns.

2. What is ChatGPT?
ChatGPT is like a super-smart text buddy created by OpenAI. It's great at chatting with people, understanding what they say, and responding in a way that makes sense. You can ask it questions, have a conversation, or get it to generate text for you. It's like having a friendly and chatty AI companion for language stuff!

3. What is Github Copilot?
GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. GitHub Copilot is built on OpenAI's Codex, which is a language model similar to the one powering this conversation (GPT-3).

GitHub Copilot is designed to assist developers by generating code suggestions as they write, helping to speed up the coding process and reduce repetitive tasks. It works as a plugin for various integrated development environments (IDEs), including Visual Studio Code. When a developer starts typing, GitHub Copilot suggests whole lines or blocks of code based on the context of the code being written.

It's important to note that the information provided here may not reflect any developments or changes that have occurred after January 2022. Please check more recent sources for the latest information on GitHub Copilot.

4. [Try it yourself: Insert your own question here]

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix

netflix <- read_csv("Data/netflix_titles.csv")
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
netflix_movies <- filter(netflix, type == "Movie")

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
count(netflix_movies, director, sort = T)
```


Which director has the most movies on netflix?
```{r}

count(netflix_movies, director, sort = T)
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

Filter the directors with the most movies but remove the NAs

```



Which director had the most movies by release year and remove the NA's?
```{r}
top_directors <- head(netflix_movies, 10)  # Adjust 10 to the desired number of top directors

# Creating a bar plot
ggplot(top_directors, aes(x = reorder(director, n), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top Directors on Netflix", x = "Director", y = "Number of Movies") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}

```



Make a bar plot of the top directors with the most movies on netflix
```{r}


```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
