---
title: "lab6warmup"
author: "Amanda Duong"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Warm-up
1. Load the `bison.csv` data.
```{r}
library("tidyverse")
library("janitor")
```
```{r}
bison <- read_csv("data/bison.csv")
```

2. What are the dimensions and structure of the data?
```{r}
dim(bison)
str(bison)
```

3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.
```{r}
names(bison)
animal <- select(bison, "animal_code", "animal_sex", "animal_weight", "animal_yob")
animal
```
```{r}
bison1 <- bison %>%
  select(animal_code, animal_sex, animal_weight, animal_yob)
```

4. Pull out the animals born between 1980-1990.
```{r}
bison2 <- bison1 %>% 
  select(animal_code, animal_sex, animal_weight, animal_yob) %>% 
  filter(between(animal_yob, 1980, 1990))
bison2
```

5. How many male and female bison are represented between 1980-1990?

```{r}
bison2 <- bison1 %>% 
  select(animal_code, animal_sex, animal_weight, animal_yob) %>% 
  filter(between(animal_yob, 1980, 1990))
  table(bison2$animal_sex)
```
```{r}
males <- bison2 %>%
  filter(animal_yob>=1980 & animal_yob <= 1990) %>%
  filter(animal_sex=="M")
males
```
```{r}
females <- bison2 %>%
  filter(animal_yob>=1980 & animal_yob <= 1990) %>%
  filter(animal_sex=="F")
females
```

6. Between 1980-1990, were males or females larger on average?
```{r}
mean(females$animal_weight)
mean(males$animal_weight)
```
Males are larger on average between 1980-1990.
